---
title: "Coastal Analysis"
output: pdf_document
---

```{r,echo=FALSE,message=FALSE,warning=FALSE}
# Set so that long lines in R will be wrapped:
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=70), tidy=TRUE)
```

## Read in data
```{r}
library(readxl)
library(lme4)
library(sjPlot)
library(sjmisc)
library(sjlabelled)

#Read in dataset with coastal coding. Read in summary sheet (sheet 13)
coastal <- read_excel("FIPS-based datasets_05232021.xlsx", sheet = 13)

#summary(coastal)

#Read in PM25 data from our 2020 study, created with: PM25 = data.frame(fips = aggregate_pm_census_cdc_test_beds$fips, mean_pm25 = aggregate_pm_census_cdc_test_beds$mean_pm25)
  #save(PM25, file = 'PM25.Rda')
load('PM25.Rda')
```

## Create smaller datasets from previous datasets, dataclean, merge region dataset with summary dataset, finally merge with PM25 dataset.
```{r}
coastal.new = data.frame(coastal$`FIPS as Text`, coastal$state, coastal$cases, coastal$deaths, coastal$`Country REGION`, coastal$`Coastal Distance`, coastal$`Population 2019 Estimate`, coastal$`Population Density`, coastal$`All Ages in Poverty (%)`, coastal$`Under 18s in Poverty`, coastal$`Median Income`, coastal$`percent adult obesity`, coastal$`diff/total`, coastal$`Politcal alignment 2020 election`, coastal$`median age 2019`, coastal$Humid)
colnames(coastal.new) = c('fips', 'state', 'cases', 'deaths', 'region', 'coastal.distance', 'population2019', 'popdensity', 'poverty', 'under18poverty', 'median_income', 'pct_obesity', 'voter_margin_2020', 'party', 'median_age', 'humidity')

#change NAs in coastal.distance to level 4, and save as factor with reference level 4.
coastal.new$coastal.distance[is.na(coastal.new$coastal.distance)] <- 4
coastal.new$coastal.distance = as.factor(coastal.new$coastal.distance)
coastal.new <- within(coastal.new, coastal.distance <- relevel(coastal.distance, ref = 4))

#change NAs in coastal region to Inland, and save as factor with reference level Inland
coastal.new$region[is.na(coastal.new$region)] <- 'Inland'
coastal.new$region[coastal.new$region == '0'] <- 'Inland'
coastal.new$region = tolower(coastal.new$region)
coastal.new$region = as.factor(coastal.new$region)
coastal.new <- within(coastal.new, region <- relevel(region, ref = 'inland'))

#Merge with PM25 dataset
coastal.new = merge(coastal.new, PM25, by = 'fips')
summary(coastal.new)
```
\newpage
## PRELIMINARY ANALYSIS on only coastal counties
```{r}
# Subset coastal counties only
coastal.only = coastal.new[coastal.new$coastal.distance != 4,]
nrow(coastal.only)
nrow(na.omit(coastal.only))

# Model cases
model.initial.cases = glmer(cases ~ (1|state) + coastal.distance + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + factor(humidity) + mean_pm25, family = poisson(link = "log"), data = coastal.only)
summary(model.initial.cases)

# Model deaths
model.initial.deaths = glmer(deaths ~ (1|state) + coastal.distance + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + factor(humidity) + mean_pm25, family = poisson(link = "log"), data = coastal.only)
summary(model.initial.deaths)
```

\newpage
## Redo: Create indicator for being a coast (levels 1,2,3) instead. 
```{r}
# Indicator Coastal or NonCoastal
coastal.new$indicatorcoast = ifelse(coastal.new$coastal.distance == '1', 'Coastal', 'NonCoastal')

# Model cases
model.indicator.cases = glmer(cases ~ (1|state) + factor(indicatorcoast) + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + factor(humidity) + mean_pm25, family = poisson(link = "log"), data = coastal.new)
summary(model.indicator.cases)

# Model deaths
model.indicator.deaths = glmer(deaths ~ (1|state) + factor(indicatorcoast) + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + factor(humidity) + mean_pm25, family = poisson(link = "log"), data = coastal.new)
summary(model.indicator.deaths)
```

## Repeat above, - humidity
```{r}
# Model cases
model.initial.cases.nohumidity = glmer(cases ~ (1|state) + coastal.distance + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25, family = poisson(link = "log"), data = coastal.only)
summary(model.initial.cases.nohumidity)

# Model deaths
model.initial.deaths.nohumidity = glmer(deaths ~ (1|state) + coastal.distance + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25, family = poisson(link = "log"), data = coastal.only)
summary(model.initial.deaths.nohumidity)
```

```{r}
# Model cases
model.indicator.cases.nohumidity = glmer(cases ~ (1|state) + factor(indicatorcoast) + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25, family = poisson(link = "log"), data = coastal.new)
summary(model.indicator.cases.nohumidity)

# Model deaths
model.indicator.deaths.nohumidity = glmer(deaths ~ (1|state) + factor(indicatorcoast) + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25, family = poisson(link = "log"), data = coastal.new)
summary(model.indicator.deaths.nohumidity)
```
\newpage
## Analysis by region
```{r}
model.byregion.cases = glmer(cases ~ (1|state) + region + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + factor(humidity) + mean_pm25, family = poisson(link = "log"), data = coastal.new)
summary(model.byregion.cases)

model.byregion.deaths = glmer(deaths ~ (1|state) + region + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + factor(humidity) + mean_pm25, family = poisson(link = "log"), data = coastal.new)
summary(model.byregion.deaths)

model.byregion.cases.nohumidity = glmer(cases ~ (1|state) + region + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25, family = poisson(link = "log"), data = coastal.new)
summary(model.byregion.cases.nohumidity)

model.byregion.deaths.nohumidity = glmer(deaths ~ (1|state) + region + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25, family = poisson(link = "log"), data = coastal.new)
summary(model.byregion.deaths.nohumidity)
```

\newpage
## Print exponentiated coefficients (ratio of probabilities)
```{r echo = F, eval = F}
paste(round(exp(summary(model.initial.cases)$coefficients[2,1]),3), " (", round(exp(summary(model.initial.cases)$coefficients[2,1] - 1.96*summary(model.initial.cases)$coefficients[2,2]),3), ", ", round(exp(summary(model.initial.cases)$coefficients[2,1] + 1.96*summary(model.initial.cases)$coefficients[2,2]),3), ")", sep = "")

paste(round(exp(summary(model.initial.cases)$coefficients[3,1]),3), " (", round(exp(summary(model.initial.cases)$coefficients[3,1] - 1.96*summary(model.initial.cases)$coefficients[3,2]),3), ", ", round(exp(summary(model.initial.cases)$coefficients[3,1] + 1.96*summary(model.initial.cases)$coefficients[3,2]),3), ")", sep = "")

paste(round(exp(summary(model.initial.deaths)$coefficients[2,1]),3), " (", round(exp(summary(model.initial.deaths)$coefficients[2,1] - 1.96*summary(model.initial.deaths)$coefficients[2,2]),3), ", ", round(exp(summary(model.initial.deaths)$coefficients[2,1] + 1.96*summary(model.initial.deaths)$coefficients[2,2]),3), ")", sep = "")

paste(round(exp(summary(model.initial.deaths)$coefficients[3,1]),3), " (", round(exp(summary(model.initial.deaths)$coefficients[3,1] - 1.96*summary(model.initial.deaths)$coefficients[3,2]),3), ", ", round(exp(summary(model.initial.deaths)$coefficients[3,1] + 1.96*summary(model.initial.deaths)$coefficients[3,2]),3), ")", sep = "")

paste(round(exp(summary(model.initial.cases.nohumidity)$coefficients[2,1]),3), " (", round(exp(summary(model.initial.cases.nohumidity)$coefficients[2,1] - 1.96*summary(model.initial.cases.nohumidity)$coefficients[2,2]),3), ", ", round(exp(summary(model.initial.cases.nohumidity)$coefficients[2,1] + 1.96*summary(model.initial.cases.nohumidity)$coefficients[2,2]),3), ")", sep = "")

paste(round(exp(summary(model.initial.cases.nohumidity)$coefficients[3,1]),3), " (", round(exp(summary(model.initial.cases.nohumidity)$coefficients[3,1] - 1.96*summary(model.initial.cases.nohumidity)$coefficients[3,2]),3), ", ", round(exp(summary(model.initial.cases.nohumidity)$coefficients[3,1] + 1.96*summary(model.initial.cases.nohumidity)$coefficients[3,2]),3), ")", sep = "")

paste(round(exp(summary(model.initial.deaths.nohumidity)$coefficients[2,1]),3), " (", round(exp(summary(model.initial.deaths.nohumidity)$coefficients[2,1] - 1.96*summary(model.initial.deaths.nohumidity)$coefficients[2,2]),3), ", ", round(exp(summary(model.initial.deaths.nohumidity)$coefficients[2,1] + 1.96*summary(model.initial.deaths.nohumidity)$coefficients[2,2]),3), ")", sep = "")

paste(round(exp(summary(model.initial.deaths.nohumidity)$coefficients[3,1]),3), " (", round(exp(summary(model.initial.deaths.nohumidity)$coefficients[3,1] - 1.96*summary(model.initial.deaths.nohumidity)$coefficients[3,2]),3), ", ", round(exp(summary(model.initial.deaths.nohumidity)$coefficients[3,1] + 1.96*summary(model.initial.deaths.nohumidity)$coefficients[3,2]),3), ")", sep = "")

paste(round(exp(summary(model.indicator.cases)$coefficients[2,1]),3), " (", round(exp(summary(model.indicator.cases)$coefficients[2,1] - 1.96*summary(model.indicator.cases)$coefficients[2,2]),3), ", ", round(exp(summary(model.indicator.cases)$coefficients[2,1] + 1.96*summary(model.indicator.cases)$coefficients[2,2]),3), ")", sep = "")

paste(round(exp(summary(model.indicator.deaths)$coefficients[2,1]),3), " (", round(exp(summary(model.indicator.deaths)$coefficients[2,1] - 1.96*summary(model.indicator.deaths)$coefficients[2,2]),3), ", ", round(exp(summary(model.indicator.deaths)$coefficients[2,1] + 1.96*summary(model.indicator.deaths)$coefficients[2,2]),3), ")", sep = "")

paste(round(exp(summary(model.indicator.cases.nohumidity)$coefficients[2,1]),3), " (", round(exp(summary(model.indicator.cases.nohumidity)$coefficients[2,1] - 1.96*summary(model.indicator.cases.nohumidity)$coefficients[2,2]),3), ", ", round(exp(summary(model.indicator.cases.nohumidity)$coefficients[2,1] + 1.96*summary(model.indicator.cases.nohumidity)$coefficients[2,2]),3), ")", sep = "")

paste(round(exp(summary(model.indicator.deaths.nohumidity)$coefficients[2,1]),3), " (", round(exp(summary(model.indicator.deaths.nohumidity)$coefficients[2,1] - 1.96*summary(model.indicator.deaths.nohumidity)$coefficients[2,2]),3), ", ", round(exp(summary(model.indicator.deaths.nohumidity)$coefficients[2,1] + 1.96*summary(model.indicator.deaths.nohumidity)$coefficients[2,2]),3), ")", sep = "")


```
```{r echo = F, eval = F}
for (i in 2:10){
  print(paste(round(exp(summary(model.byregion.cases)$coefficients[i,1]),3), " (", ... = round(exp(summary(model.byregion.cases)$coefficients[i,1] - 1.96*summary(model.byregion.cases)$coefficients[i,2]),3), ", ", round(exp(summary(model.byregion.cases)$coefficients[i,1] + 1.96*summary(model.byregion.cases)$coefficients[i,2]),3), ")", sep = ""))

  print(paste(round(exp(summary(model.byregion.deaths)$coefficients[i,1]),3), " (", round(exp(summary(model.byregion.deaths)$coefficients[i,1] - 1.96*summary(model.byregion.deaths)$coefficients[i,2]),3), ", ", round(exp(summary(model.byregion.deaths)$coefficients[i,1] + 1.96*summary(model.byregion.deaths)$coefficients[i,2]),3), ")", sep = ""))

  print(paste(round(exp(summary(model.byregion.cases.nohumidity)$coefficients[i,1]),3), " (", round(exp(summary(model.byregion.cases.nohumidity)$coefficients[i,1] - 1.96*summary(model.byregion.cases.nohumidity)$coefficients[i,2]),3), ", ", round(exp(summary(model.byregion.cases.nohumidity)$coefficients[i,1] + 1.96*summary(model.byregion.cases.nohumidity)$coefficients[i,2]),3), ")", sep = ""))

  print(paste(round(exp(summary(model.byregion.deaths.nohumidity)$coefficients[i,1]),3), " (", round(exp(summary(model.byregion.deaths.nohumidity)$coefficients[i,1] - 1.96*summary(model.byregion.deaths.nohumidity)$coefficients[i,2]),3), ", ", round(exp(summary(model.byregion.deaths.nohumidity)$coefficients[i,1] + 1.96*summary(model.byregion.deaths.nohumidity)$coefficients[i,2]),3), ")", sep = ""))
}
```

## Printing tables
```{r eval = F}
tab_model(model.initial.cases, digits = 3)
tab_model(model.initial.deaths, digits = 3)
tab_model(model.initial.cases.nohumidity, digits = 3)
tab_model(model.initial.deaths.nohumidity, digits = 3)
tab_model(model.indicator.cases, digits = 3)
tab_model(model.indicator.deaths, digits = 3)
tab_model(model.indicator.cases.nohumidity, digits = 3)
tab_model(model.indicator.deaths.nohumidity, digits = 3)
tab_model(model.byregion.cases, digits = 3)
tab_model(model.byregion.deaths, digits = 3)
tab_model(model.byregion.cases.nohumidity, digits = 3)
tab_model(model.byregion.deaths.nohumidity, digits = 3)
```












