---
title: "Coastal Analysis"
output: pdf_document
---

```{r,echo=FALSE,message=FALSE,warning=FALSE}
# Set so that long lines in R will be wrapped:
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=70), tidy=TRUE)
```
## TLDR: To replicate tables in slides, run tab_model commands at the bottom (that is, after running all the models)

## Read in data
```{r}
library(readxl)
library(lme4)
library(sjPlot)
library(sjmisc)
library(sjlabelled)

#Read in dataset with coastal coding. Read in summary sheet (sheet 13)
coastal <- read_excel("FIPS-based datasets_05232021.xlsx", sheet = 13)

#summary(coastal)

#Read in PM25 and humidity data from our 2020 study, created with: confounding = data.frame(fips = aggregate_pm_census_cdc_test_beds$fips, q_popdensity = aggregate_pm_census_cdc_test_beds$q_popdensity, poverty.old = aggregate_pm_census_cdc_test_beds$poverty, median_house_value = aggregate_pm_census_cdc_test_beds$median_house_value, median_household_income = aggregate_pm_census_cdc_test_beds$median_household_income, owner_occupied = aggregate_pm_census_cdc_test_beds$owner_occupied, blk_pct = aggregate_pm_census_cdc_test_beds$blk_pct, hispanic_pct = aggregate_pm_census_cdc_test_beds$hispanic_pct, white_pct = aggregate_pm_census_cdc_test_beds$white_pct, native_pct = aggregate_pm_census_cdc_test_beds$native_pct, asian_pct = aggregate_pm_census_cdc_test_beds$asian_pct, no_grad = aggregate_pm_census_cdc_test_beds$no_grad, date_since_social =  aggregate_pm_census_cdc_test_beds$date_since_social, date_since = aggregate_pm_census_cdc_test_beds$date_since, beds = aggregate_pm_census_cdc_test_beds$beds, population.old = aggregate_pm_census_cdc_test_beds$population, obese = aggregate_pm_census_cdc_test_beds$obese, smoke = aggregate_pm_census_cdc_test_beds$smoke, mean_summer_temp = aggregate_pm_census_cdc_test_beds$mean_summer_temp, mean_winter_temp =  aggregate_pm_census_cdc_test_beds$mean_winter_temp, mean_pm25 = aggregate_pm_census_cdc_test_beds$mean_pm25, mean_summer_rm = aggregate_pm_census_cdc_test_beds$mean_summer_rm, mean_winter_rm = aggregate_pm_census_cdc_test_beds$mean_winter_rm)
  #save(confounding, file = 'confounding.Rda')
load('confounding.Rda')
```

## Create smaller datasets from previous datasets, dataclean, merge region dataset with summary dataset, finally merge with PM25 dataset.
```{r}
coastal.new = data.frame(coastal$`FIPS as Text`, coastal$state, coastal$cases, coastal$deaths, coastal$`Country REGION`, coastal$`Coastal Distance`, coastal$`Population 2019 Estimate`, coastal$`Population Density`, coastal$`All Ages in Poverty (%)`, coastal$`Under 18s in Poverty`, coastal$`Median Income`, coastal$`percent adult obesity`, coastal$`diff/total`, coastal$`Politcal alignment 2020 election`, coastal$`median age 2019`, coastal$Humid)
colnames(coastal.new) = c('fips', 'state', 'cases', 'deaths', 'region', 'coastal.distance', 'population2019', 'popdensity', 'poverty', 'under18poverty', 'median_income', 'pct_obesity', 'voter_margin_2020', 'party', 'median_age', 'humidity')

#change NAs in coastal.distance to level 4, and save as factor with reference level 4.
coastal.new$coastal.distance[is.na(coastal.new$coastal.distance)] <- 4
coastal.new$coastal.distance = as.factor(coastal.new$coastal.distance)
coastal.new <- within(coastal.new, coastal.distance <- relevel(coastal.distance, ref = 4))

#change NAs in coastal region to Inland, and save as factor with reference level Inland
coastal.new$region[is.na(coastal.new$region)] <- 'Inland'
coastal.new$region[coastal.new$region == '0'] <- 'Inland'
coastal.new$region = tolower(coastal.new$region)
coastal.new$region = as.factor(coastal.new$region)
coastal.new <- within(coastal.new, region <- relevel(region, ref = 'inland'))

#Merge with confounding dataset
coastal.new = merge(coastal.new, confounding, by = 'fips')
names(coastal.new)[names(coastal.new) == 'poverty.x'] <- 'poverty'
summary(coastal.new)
```
\newpage
## PRELIMINARY ANALYSIS on only coastal counties BEING EDITED
```{r echo = F, eval = F}
# Subset coastal counties only
coastal.only = coastal.new[coastal.new$coastal.distance != 4,]
nrow(coastal.only)
nrow(na.omit(coastal.only))

# Model cases
model.initial.cases = glmer(cases ~ (1|state) + coastal.distance + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + factor(humidity) + mean_pm25, family = poisson(link = "log"), data = coastal.only)
summary(model.initial.cases)

# Model deaths
model.initial.deaths = glmer(deaths ~ (1|state) + coastal.distance + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + factor(humidity) + mean_pm25, family = poisson(link = "log"), data = coastal.only)
summary(model.initial.deaths)
```

## Create indicator for being a coast (degree 1) 
```{r}
# Indicator Coastal or NonCoastal
coastal.new$indicatorcoast = ifelse(coastal.new$coastal.distance == '1', 'Coastal', 'NonCoastal')

# Model cases
model.indicator.cases = glmer(cases ~ (1|state) + factor(indicatorcoast) + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25 + mean_summer_rm + mean_winter_rm, family = poisson(link = "log"), data = coastal.new)
summary(model.indicator.cases)

# Model deaths
model.indicator.deaths = glmer(deaths ~ (1|state) + factor(indicatorcoast) + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25 + mean_summer_rm + mean_winter_rm, family = poisson(link = "log"), data = coastal.new)
summary(model.indicator.deaths)
```

## Repeat above, - humidity
```{r eval = F}
# being edited
# Model cases
model.initial.cases.nohumidity = glmer(cases ~ (1|state) + coastal.distance + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25, family = poisson(link = "log"), data = coastal.only)
summary(model.initial.cases.nohumidity)

# Model deaths
model.initial.deaths.nohumidity = glmer(deaths ~ (1|state) + coastal.distance + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25, family = poisson(link = "log"), data = coastal.only)
summary(model.initial.deaths.nohumidity)
```

```{r}
# Model cases
model.indicator.cases.nohumidity = glmer(cases ~ (1|state) + factor(indicatorcoast) + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25, family = poisson(link = "log"), data = coastal.new)
summary(model.indicator.cases.nohumidity)

# Model deaths
model.indicator.deaths.nohumidity = glmer(deaths ~ (1|state) + factor(indicatorcoast) + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25, family = poisson(link = "log"), data = coastal.new)
summary(model.indicator.deaths.nohumidity)
```
\newpage
## Analysis by region
```{r}
model.byregion.cases = glmer(cases ~ (1|state) + region + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25 + mean_summer_rm + mean_winter_rm, family = poisson(link = "log"), data = coastal.new)
summary(model.byregion.cases)

model.byregion.deaths = glmer(deaths ~ (1|state) + region + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25 + mean_summer_rm + mean_winter_rm, family = poisson(link = "log"), data = coastal.new)
summary(model.byregion.deaths)

model.byregion.cases.nohumidity = glmer(cases ~ (1|state) + region + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25, family = poisson(link = "log"), data = coastal.new)
summary(model.byregion.cases.nohumidity)

model.byregion.deaths.nohumidity = glmer(deaths ~ (1|state) + region + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25, family = poisson(link = "log"), data = coastal.new)
summary(model.byregion.deaths.nohumidity)
```
## By region, splitting into Urban and Rural
```{r}
coastal.new$area = ifelse(coastal.new$popdensity >= 1500, 'Urban', 'Rural')
summary(as.factor(coastal.new$area))

coastal.new$regionru = paste(as.character(coastal.new$region), coastal.new$area)
coastal.new$regionru[coastal.new$regionru == 'inland Rural'] = 'inland'
coastal.new$regionru[coastal.new$regionru == 'inland Urban'] = 'inland'
coastal.new$regionru = as.factor(coastal.new$regionru)
coastal.new <- within(coastal.new, regionru <- relevel(regionru, ref = 'inland'))
summary(coastal.new$regionru)

model.byregionru.cases = glmer(cases ~ (1|state) + regionru + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25 + mean_summer_rm + mean_winter_rm, family = poisson(link = "log"), data = coastal.new)
summary(model.byregionru.cases)

model.byregionru.deaths = glmer(deaths ~ (1|state) + regionru + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25 + mean_summer_rm + mean_winter_rm, family = poisson(link = "log"), data = coastal.new)
summary(model.byregionru.deaths)

model.byregionru.cases.nohumidity = glmer(cases ~ (1|state) + regionru + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25, family = poisson(link = "log"), data = coastal.new)
summary(model.byregionru.cases.nohumidity)

model.byregionru.deaths.nohumidity = glmer(deaths ~ (1|state) + regionru + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25, family = poisson(link = "log"), data = coastal.new)
summary(model.byregionru.deaths.nohumidity)
```



\newpage
## Print tables
```{r eval = F}
# Run each line individually not all at once
tab_model(model.indicator.cases, digits = 3)
tab_model(model.indicator.deaths, digits = 3)
tab_model(model.indicator.cases.nohumidity, digits = 3)
tab_model(model.indicator.deaths.nohumidity, digits = 3)

tab_model(model.byregion.cases, digits = 3)
tab_model(model.byregion.deaths, digits = 3)
tab_model(model.byregion.cases.nohumidity, digits = 3)
tab_model(model.byregion.deaths.nohumidity, digits = 3)

tab_model(model.byregionru.cases, digits = 3)
tab_model(model.byregionru.deaths, digits = 3)
tab_model(model.byregionru.cases.nohumidity, digits = 3)
tab_model(model.byregionru.deaths.nohumidity, digits = 3)
```

## Same Analysis with Our Additional Confounders
```{r}
model.indicator.cases.addconfounders = glmer(cases ~ (1|state) + factor(indicatorcoast) + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) 
                              + scale(log(median_house_value)) + scale(owner_occupied) 
                              + scale(blk_pct) + scale(hispanic_pct) 
                              + scale(native_pct) + scale(asian_pct) 
                              + scale(date_since_social) + scale(date_since) 
                              + scale(beds/population.old) + scale(smoke) 
                              + scale(mean_pm25) 
                              + scale(mean_summer_rm) + scale(mean_winter_rm) 
                              + scale(mean_summer_temp) + scale(mean_winter_temp)
                              + scale(no_grad), family = poisson(link = "log"), data = coastal.new)

model.indicator.deaths.addconfounders = glmer(deaths ~ (1|state) + factor(indicatorcoast) + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) 
                              + scale(log(median_house_value)) + scale(owner_occupied) 
                              + scale(blk_pct) + scale(hispanic_pct)
                              + scale(native_pct) + scale(asian_pct) 
                              + scale(date_since_social) + scale(date_since) 
                              + scale(beds/population.old) + scale(smoke) 
                              + scale(mean_pm25) 
                              + scale(mean_summer_rm) + scale(mean_winter_rm) 
                              + scale(mean_summer_temp) + scale(mean_winter_temp)
                              + scale(no_grad), family = poisson(link = "log"), data = coastal.new)

model.byregion.cases.addconfounders = glmer(cases ~ (1|state) + region + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) 
                              + scale(log(median_house_value)) + scale(owner_occupied) 
                              + scale(blk_pct) + scale(hispanic_pct)
                              + scale(native_pct) + scale(asian_pct) 
                              + scale(date_since_social) + scale(date_since) 
                              + scale(beds/population.old) + scale(smoke) 
                              + scale(mean_pm25) 
                              + scale(mean_summer_rm) + scale(mean_winter_rm) 
                              + scale(mean_summer_temp) + scale(mean_winter_temp)
                              + scale(no_grad), family = poisson(link = "log"), data = coastal.new)

model.byregion.deaths.addconfounders = glmer(deaths ~ (1|state) + region + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) 
                              + scale(log(median_house_value)) + scale(owner_occupied) 
                              + scale(blk_pct) + scale(hispanic_pct) 
                              + scale(native_pct) + scale(asian_pct) 
                              + scale(date_since_social) + scale(date_since) 
                              + scale(beds/population.old) + scale(smoke) 
                              + scale(mean_pm25) 
                              + scale(mean_summer_rm) + scale(mean_winter_rm) 
                              + scale(mean_summer_temp) + scale(mean_winter_temp)
                              + scale(no_grad), family = poisson(link = "log"), data = coastal.new)

# - humidity
model.indicator.cases.addconfounders.nohumidity = glmer(cases ~ (1|state) + factor(indicatorcoast) + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) 
                              + scale(log(median_house_value)) + scale(owner_occupied) 
                              + scale(blk_pct) + scale(hispanic_pct) 
                              + scale(native_pct) + scale(asian_pct) 
                              + scale(date_since_social) + scale(date_since) 
                              + scale(beds/population.old) + scale(smoke) 
                              + scale(mean_pm25) 
                              #+ scale(mean_summer_rm) + scale(mean_winter_rm) 
                              + scale(mean_summer_temp) + scale(mean_winter_temp)
                              + scale(no_grad), family = poisson(link = "log"), data = coastal.new)

model.indicator.deaths.addconfounders.nohumidity = glmer(deaths ~ (1|state) + factor(indicatorcoast) + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) 
                              + scale(log(median_house_value)) + scale(owner_occupied) 
                              + scale(blk_pct) + scale(hispanic_pct) 
                              + scale(native_pct) + scale(asian_pct) 
                              + scale(date_since_social) + scale(date_since) 
                              + scale(beds/population.old) + scale(smoke) 
                              + scale(mean_pm25) 
                              #+ scale(mean_summer_rm) + scale(mean_winter_rm) 
                              + scale(mean_summer_temp) + scale(mean_winter_temp)
                              + scale(no_grad), family = poisson(link = "log"), data = coastal.new)

model.byregion.cases.addconfounders.nohumidity = glmer(cases ~ (1|state) + region + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) 
                              + scale(log(median_house_value)) + scale(owner_occupied) 
                              + scale(blk_pct) + scale(hispanic_pct) 
                              + scale(native_pct) + scale(asian_pct) 
                              + scale(date_since_social) + scale(date_since) 
                              + scale(beds/population.old) + scale(smoke) 
                              + scale(mean_pm25) 
                              #+ scale(mean_summer_rm) + scale(mean_winter_rm) 
                              + scale(mean_summer_temp) + scale(mean_winter_temp)
                              + scale(no_grad), family = poisson(link = "log"), data = coastal.new)

model.byregion.deaths.addconfounders.nohumidity = glmer(deaths ~ (1|state) + region + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) 
                              + scale(log(median_house_value)) + scale(owner_occupied) 
                              + scale(blk_pct) + scale(hispanic_pct) 
                              + scale(native_pct) + scale(asian_pct)
                              + scale(date_since_social) + scale(date_since) 
                              + scale(beds/population.old) + scale(smoke) 
                              + scale(mean_pm25) 
                              #+ scale(mean_summer_rm) + scale(mean_winter_rm) 
                              + scale(mean_summer_temp) + scale(mean_winter_temp)
                              + scale(no_grad), family = poisson(link = "log"), data = coastal.new)

# Analysis by region, rural/urban split
model.byregionru.cases.addconfounders = glmer(cases ~ (1|state) + regionru + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) 
                              + scale(log(median_house_value)) + scale(owner_occupied) 
                              + scale(blk_pct) + scale(hispanic_pct)
                              + scale(native_pct) + scale(asian_pct) 
                              + scale(date_since_social) + scale(date_since) 
                              + scale(beds/population.old) + scale(smoke) 
                              + scale(mean_pm25) 
                              + scale(mean_summer_rm) + scale(mean_winter_rm) 
                              + scale(mean_summer_temp) + scale(mean_winter_temp)
                              + scale(no_grad), family = poisson(link = "log"), data = coastal.new)

model.byregionru.deaths.addconfounders = glmer(deaths ~ (1|state) + regionru + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) 
                              + scale(log(median_house_value)) + scale(owner_occupied) 
                              + scale(blk_pct) + scale(hispanic_pct) 
                              + scale(native_pct) + scale(asian_pct) 
                              + scale(date_since_social) + scale(date_since) 
                              + scale(beds/population.old) + scale(smoke) 
                              + scale(mean_pm25) 
                              + scale(mean_summer_rm) + scale(mean_winter_rm) 
                              + scale(mean_summer_temp) + scale(mean_winter_temp)
                              + scale(no_grad), family = poisson(link = "log"), data = coastal.new)

model.byregionru.cases.addconfounders.nohumidity = glmer(cases ~ (1|state) + regionru + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) 
                              + scale(log(median_house_value)) + scale(owner_occupied) 
                              + scale(blk_pct) + scale(hispanic_pct) 
                              + scale(native_pct) + scale(asian_pct) 
                              + scale(date_since_social) + scale(date_since) 
                              + scale(beds/population.old) + scale(smoke) 
                              + scale(mean_pm25) 
                              #+ scale(mean_summer_rm) + scale(mean_winter_rm) 
                              + scale(mean_summer_temp) + scale(mean_winter_temp)
                              + scale(no_grad), family = poisson(link = "log"), data = coastal.new)

model.byregionru.deaths.addconfounders.nohumidity = glmer(deaths ~ (1|state) + regionru + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) 
                              + scale(log(median_house_value)) + scale(owner_occupied) 
                              + scale(blk_pct) + scale(hispanic_pct) 
                              + scale(native_pct) + scale(asian_pct)
                              + scale(date_since_social) + scale(date_since) 
                              + scale(beds/population.old) + scale(smoke) 
                              + scale(mean_pm25) 
                              #+ scale(mean_summer_rm) + scale(mean_winter_rm) 
                              + scale(mean_summer_temp) + scale(mean_winter_temp)
                              + scale(no_grad), family = poisson(link = "log"), data = coastal.new)
```

## Printing tables
```{r eval = F}
# Run each line individually not all at once
tab_model(model.indicator.cases.addconfounders, digits = 3)
tab_model(model.indicator.cases.addconfounders.nohumidity, digits = 3)
tab_model(model.indicator.deaths.addconfounders, digits = 3)
tab_model(model.indicator.deaths.addconfounders.nohumidity, digits = 3)

tab_model(model.byregion.cases.addconfounders, digits = 3)
tab_model(model.byregion.cases.addconfounders.nohumidity, digits = 3)
tab_model(model.byregion.deaths.addconfounders, digits = 3)
tab_model(model.byregion.deaths.addconfounders.nohumidity, digits = 3)

tab_model(model.byregionru.cases.addconfounders, digits = 3)
tab_model(model.byregionru.cases.addconfounders.nohumidity, digits = 3)
tab_model(model.byregionru.deaths.addconfounders, digits = 3)
tab_model(model.byregionru.deaths.addconfounders.nohumidity, digits = 3)
```















