---
title: "Coastal Analysis"
output: pdf_document
---

```{r,echo=FALSE,message=FALSE,warning=FALSE}
# Set so that long lines in R will be wrapped:
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

## Read in data
```{r}
library("readxl")
library("lme4")

#Read in dataset with coastal coding. Read in summary sheet (sheet 13)
coastal <- read_excel("FIPS-based datasets_05232021.xlsx", sheet = 13)
#summary(coastal)

#Read in PM25 data from our 2020 study, created with: PM25 = data.frame(fips = aggregate_pm_census_cdc_test_beds$fips, mean_pm25 = aggregate_pm_census_cdc_test_beds$mean_pm25)
  #save(PM25, file = 'PM25.Rda')
load('PM25.Rda')
```

## Create smaller dataset from previous dataset, dataclean, merge with PM25 dataset.
```{r}
coastal.new = data.frame(coastal$`FIPS as Text`, coastal$state, coastal$cases, coastal$deaths, coastal$`Country REGION`, coastal$`Coastal Distance`, coastal$`Population 2019 Estimate`, coastal$`Population Density`, coastal$`All Ages in Poverty (%)`, coastal$`Under 18s in Poverty`, coastal$`Median Income`, coastal$`percent adult obesity`, coastal$`diff/total`, coastal$`Politcal alignment 2020 election`, coastal$`median age 2019`, coastal$Humid)
colnames(coastal.new) = c('fips', 'state', 'cases', 'deaths', 'region', 'coastal.distance', 'population2019', 'popdensity', 'poverty', 'under18poverty', 'median_income', 'pct_obesity', 'voter_margin_2020', 'party', 'median_age', 'humidity')

#change NAs in coastal.distance to level 4, and save as factor with reference level 4.
coastal.new$coastal.distance[is.na(coastal.new$coastal.distance)] <- 4
coastal.new$coastal.distance = as.factor(coastal.new$coastal.distance)
coastal.new <- within(coastal.new, coastal.distance <- relevel(coastal.distance, ref = 4))

#change NAs in 'region' to 'Inland', convert all characters to lowercase
coastal.new$region[is.na(coastal.new$region)] <- 'Inland'
coastal.new$region = tolower(coastal.new$region)

#Merge with PM25 dataset
coastal.new = merge(coastal.new, PM25, by = 'fips')
summary(coastal.new)
```
\newpage
## PRELIMINARY ANALYSIS on only coastal counties
```{r}
# Subset coastal counties only
coastal.only = coastal.new[coastal.new$coastal.distance != 4,]
nrow(coastal.only)
nrow(na.omit(coastal.only))

# Model cases
model.initial.cases = glmer(cases ~ (1|state) + coastal.distance + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + factor(humidity) + mean_pm25, family = poisson(link = "log"), data = coastal.only)
summary(model.initial.cases)

# Model deaths
model.initial.deaths = glmer(deaths ~ (1|state) + coastal.distance + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + factor(humidity) + mean_pm25, family = poisson(link = "log"), data = coastal.only)
summary(model.initial.deaths)
```

\newpage
## Redo: Create indicator for being a coast (levels 1,2,3) instead. 
```{r}
# Indicator Coastal or NonCoastal
coastal.new$indicatorcoast = ifelse(coastal.new$coastal.distance == '4', 'Noncoastal', 'Coastal')

# Model cases
model.indicator.cases = glmer(cases ~ (1|state) + factor(indicatorcoast) + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + factor(humidity) + mean_pm25, family = poisson(link = "log"), data = coastal.new)
summary(model.indicator.cases)

# Model deaths
model.indicator.deaths = glmer(deaths ~ (1|state) + factor(indicatorcoast) + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + factor(humidity) + mean_pm25, family = poisson(link = "log"), data = coastal.new)
summary(model.indicator.deaths)
```

## Repeat above, - humidity
```{r}
# Model cases
model.initial.cases.nohumidity = glmer(cases ~ (1|state) + coastal.distance + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25, family = poisson(link = "log"), data = coastal.only)
summary(model.initial.cases.nohumidity)

# Model deaths
model.initial.deaths.nohumidity = glmer(deaths ~ (1|state) + coastal.distance + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25, family = poisson(link = "log"), data = coastal.only)
summary(model.initial.deaths.nohumidity)
```

```{r}
# Model cases
model.indicator.cases.nohumidity = glmer(cases ~ (1|state) + factor(indicatorcoast) + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25, family = poisson(link = "log"), data = coastal.new)
summary(model.indicator.cases.nohumidity)

# Model deaths
model.indicator.deaths.nohumidity = glmer(deaths ~ (1|state) + factor(indicatorcoast) + offset(log(population2019)) + scale(popdensity) + scale(poverty) + scale(log(median_income)) + scale(pct_obesity) + scale(voter_margin_2020) + scale(median_age) + factor(party) + mean_pm25, family = poisson(link = "log"), data = coastal.new)
summary(model.indicator.deaths.nohumidity)
```

## Summary of output
```{r echo = F}
paste(signif(summary(model.initial.cases)$coefficients[2,1],2), " (", signif(summary(model.initial.cases)$coefficients[2,1] - 1.96*summary(model.initial.cases)$coefficients[2,2],2), ", ", signif(summary(model.initial.cases)$coefficients[2,1] + 1.96*summary(model.initial.cases)$coefficients[2,2],2), ")", sep = "")

paste(signif(summary(model.initial.cases)$coefficients[3,1],2), " (", signif(summary(model.initial.cases)$coefficients[3,1] - 1.96*summary(model.initial.cases)$coefficients[3,2],2), ", ", signif(summary(model.initial.cases)$coefficients[3,1] + 1.96*summary(model.initial.cases)$coefficients[3,2],2), ")", sep = "")

paste(signif(summary(model.initial.deaths)$coefficients[2,1],2), " (", signif(summary(model.initial.deaths)$coefficients[2,1] - 1.96*summary(model.initial.deaths)$coefficients[2,2],2), ", ", signif(summary(model.initial.deaths)$coefficients[2,1] + 1.96*summary(model.initial.deaths)$coefficients[2,2],2), ")", sep = "")

paste(signif(summary(model.initial.deaths)$coefficients[3,1],2), " (", signif(summary(model.initial.deaths)$coefficients[3,1] - 1.96*summary(model.initial.deaths)$coefficients[3,2],2), ", ", signif(summary(model.initial.deaths)$coefficients[3,1] + 1.96*summary(model.initial.deaths)$coefficients[3,2],2), ")", sep = "")

paste(signif(summary(model.initial.cases.nohumidity)$coefficients[2,1],2), " (", signif(summary(model.initial.cases.nohumidity)$coefficients[2,1] - 1.96*summary(model.initial.cases.nohumidity)$coefficients[2,2],2), ", ", signif(summary(model.initial.cases.nohumidity)$coefficients[2,1] + 1.96*summary(model.initial.cases.nohumidity)$coefficients[2,2],2), ")", sep = "")

paste(signif(summary(model.initial.cases.nohumidity)$coefficients[3,1],2), " (", signif(summary(model.initial.cases.nohumidity)$coefficients[3,1] - 1.96*summary(model.initial.cases.nohumidity)$coefficients[3,2],2), ", ", signif(summary(model.initial.cases.nohumidity)$coefficients[3,1] + 1.96*summary(model.initial.cases.nohumidity)$coefficients[3,2],2), ")", sep = "")

paste(signif(summary(model.initial.deaths.nohumidity)$coefficients[2,1],2), " (", signif(summary(model.initial.deaths.nohumidity)$coefficients[2,1] - 1.96*summary(model.initial.deaths.nohumidity)$coefficients[2,2],2), ", ", signif(summary(model.initial.deaths.nohumidity)$coefficients[2,1] + 1.96*summary(model.initial.deaths.nohumidity)$coefficients[2,2],2), ")", sep = "")

paste(signif(summary(model.initial.deaths.nohumidity)$coefficients[3,1],2), " (", signif(summary(model.initial.deaths.nohumidity)$coefficients[3,1] - 1.96*summary(model.initial.deaths.nohumidity)$coefficients[3,2],2), ", ", signif(summary(model.initial.deaths.nohumidity)$coefficients[3,1] + 1.96*summary(model.initial.deaths.nohumidity)$coefficients[3,2],2), ")", sep = "")

paste(signif(summary(model.indicator.cases)$coefficients[2,1],2), " (", signif(summary(model.indicator.cases)$coefficients[2,1] - 1.96*summary(model.indicator.cases)$coefficients[2,2],2), ", ", signif(summary(model.indicator.cases)$coefficients[2,1] + 1.96*summary(model.indicator.cases)$coefficients[2,2],2), ")", sep = "")


paste(signif(summary(model.indicator.deaths)$coefficients[2,1],2), " (", signif(summary(model.indicator.deaths)$coefficients[2,1] - 1.96*summary(model.indicator.deaths)$coefficients[2,2],2), ", ", signif(summary(model.indicator.deaths)$coefficients[2,1] + 1.96*summary(model.indicator.deaths)$coefficients[2,2],2), ")", sep = "")

paste(signif(summary(model.indicator.cases.nohumidity)$coefficients[2,1],2), " (", signif(summary(model.indicator.cases.nohumidity)$coefficients[2,1] - 1.96*summary(model.indicator.cases.nohumidity)$coefficients[2,2],2), ", ", signif(summary(model.indicator.cases.nohumidity)$coefficients[2,1] + 1.96*summary(model.indicator.cases.nohumidity)$coefficients[2,2],2), ")", sep = "")

paste(signif(summary(model.indicator.deaths.nohumidity)$coefficients[2,1],2), " (", signif(summary(model.indicator.deaths.nohumidity)$coefficients[2,1] - 1.96*summary(model.indicator.deaths.nohumidity)$coefficients[2,2],2), ", ", signif(summary(model.indicator.deaths.nohumidity)$coefficients[2,1] + 1.96*summary(model.indicator.deaths.nohumidity)$coefficients[2,2],2), ")", sep = "")
```
\begin{center}
\begin{tabular}{ c| c | c }
 \multicolumn{3}{c}{\textbf{ Humidity Included}} \\
 \hline
 Coastal Distance & Case Model Coefficients (CI) & Death Model Coefficients (CI) \\
 \hline
 1 (Reference) &  0  &  0 \\ 
 2 & -0.0083 (-0.0099, -0.0068) & 0.042 (0.031, 0.053)\\  
 3 & 0.0033 (0.0013, 0.0053) & 0.024 (0.0091, 0.038)
\end{tabular}
\end{center}

\begin{center}
\begin{tabular}{ c| c | c }
 \multicolumn{3}{c}{\textbf{ Humidity Excluded}} \\
 \hline
 Coastal Distance & Case Model Coefficients (CI) & Death Model Coefficients (CI) \\
 \hline
 1 (Reference) & 0  &  0 \\ 
 2 & 0.013 (0.011, 0.014) & 0.058 (0.047, 0.069)\\  
 3 & 0.022 (0.02, 0.024) & 0.036 (0.022, 0.05)
\end{tabular}
\end{center}

\begin{center}
\begin{tabular}{ c| c | c }
 \multicolumn{3}{c}{\textbf{ Humidity Included}} \\
 \hline
 Binary Variable & Case Model Coefficients (CI) & Death Model Coefficients (CI) \\
 \hline
 Coastal (Reference) &  0  &  0 \\ 
 Noncoastal & 0.0018 (0.00059, 0.003) & -0.039 (-0.048, -0.03)
\end{tabular}
\end{center}

\begin{center}
\begin{tabular}{ c| c | c }
 \multicolumn{3}{c}{\textbf{ Humidity Excluded}} \\
 \hline
 Binary Variable & Case Model Coefficients (CI) & Death Model Coefficients (CI) \\
 \hline
 Coastal (Reference) &  0  &  0 \\ 
 Noncoastal & 0.023 (0.022, 0.024) & -0.0085 (-0.017, 1e-04)
\end{tabular}
\end{center}





